---
title: "STAT 2210 - Homework 4"
author: "Giang Vu"
date: "10/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
options(scipen = 999)
setwd("/Users/giangvu/Desktop/STAT 2210 - Categorical/HW")
```

## **1. Problem 1**

### **(a) Equation for logistic regression model**

$$ln(\frac{\hat{\pi}}{1-\hat{\pi}})=-3.77714+0.14486(LI)$$

### **(b) Estimated probabilities of remission**

When $LI=8$,\
$$\hat{\pi}(8)=\frac{e^{-3.77714+0.14486(8)}}{1+e^{-3.77714+0.14486(8)}} = 0.06797244$$

When $LI=26$,\
$$\hat{\pi}(26)=\frac{e^{-3.77714+0.14486(26)}}{1+e^{-3.77714+0.14486(26)}} = 0.497305$$

When $LI=34$,\
$$\hat{\pi}(34)=\frac{e^{-3.77714+0.14486(34)}}{1+e^{-3.77714+0.14486(34)}} = 0.7591637$$

```{r include=FALSE}
exp(-3.77714+0.14486*8)/(1+exp(-3.77714+0.14486*8))
exp(-3.77714+0.14486*26)/(1+exp(-3.77714+0.14486*26))
exp(-3.77714+0.14486*34)/(1+exp(-3.77714+0.14486*34))
```

### **(c) Rate of change in $\hat\pi$ when $LI=26$**
$$\hat\beta_1\hat\pi(26)(1-\hat\pi(26))=0.14486(0.497305)(1-0.497305)=0.036$$
```{r include=F}
0.14486*(0.497305)*(1-0.497305)
```

### **(d) Change in $\hat\pi$ from $LI=14$ to $LI=28$**

When $LI=14$,\
$$\hat{\pi}(14)=\frac{e^{-3.77714+0.14486(14)}}{1+e^{-3.77714+0.14486(14)}} = 0.148$$
When $LI=28$,\
$$\hat{\pi}(28)=\frac{e^{-3.77714+0.14486(28)}}{1+e^{-3.77714+0.14486(28)}} = 0.569$$
Therefore,\
$$\hat{\pi}(28)-\hat{\pi}(14)=0.57-0.15=0.42$$

```{r include=FALSE}
exp(-3.77714+0.14486*14)/(1+exp(-3.77714+0.14486*14))
exp(-3.77714+0.14486*28)/(1+exp(-3.77714+0.14486*28))
```

### **(e) Change in estimated odds of remission with a one unit increase in LI**
$$e^{\hat\beta_1}=e^{0.14486}=1.16$$
```{r include=F}
exp(0.14486)
```

### **(f) 95% confidence interval for $\pi$ of remission when $LI=26$**
The 95% CI for $\beta_0 + \beta_1 x$ when $LI=26$ is
$$\hat\pi(26) \pm 1.96 \sqrt{Var(\hat\beta_0 + \hat\beta_1 (26))}$$
We have \
$$Var(\hat\beta_0 + \hat\beta_1 (26)) = Var(\hat\beta_0) + 2(26)Cov(\hat\beta_0, \hat\beta_1) + 26^2 Var(\hat\beta_1)= 1.9 - 2(26)(0.077)+26^2(0.004)=0.6$$

```{r include=F}
1.9 - 2*(26)*(0.077)+26^2*(0.004)
```
Therefore, the 95% CI for $\beta_0 + \beta_1 x$ when $LI=26$ is \ 
$$0.497305 \pm 1.96 \sqrt{0.6} $$
which is $(-1.021, 2.016)$. \
So, the 95% for $\pi$ when $LI=26$ is $(0.265, 0.882)$. \

### **(g) Wald test for the LI effect at $\alpha=0.05$**
Null hypothesis: $H_0: \beta_1=0$\
Alternative hypothesis: $H_A: \beta_1\neq0$\
Test statistic is\
$$Z = \frac{\hat\beta_1}{SE(\hat\beta_1)} = \frac{0.145}{0.06}=2.417$$
which follows the standard normal distribution. Thus the p-value is\
$$2P(Z \geq2.417)=2(1-0.992)=0.015 < \alpha$$
Therefore, we reject the null hypothesis and conclude that we have strong evidence of a positive effect of LI on the cancer remission probability.\
```{r include=F}
(1-pnorm(2.417))*2
```

### **(h) 95% Wald CI for the odds ratio corresponding to a 1-unit increase in LI**
The odds ratio is $e^{\beta_1}$, so we need to construct 95% Wald CI for log odds ratio first, which is $\beta_1$.\
The 95% Wald CI for $beta_1$ is $0.145 \pm 1.96(0.06)$ which will be $(0.0274, 0.2626)$.\
Therefore, the 95% Wald CI for the odds ratio is $(1.03, 1.30)$.\
*Interpretation:* We are 95% confident that a 1-unit increase in LI has at least 3% increase and at most 30% increase in the odds that a patient achieves cancer remission.\

### **(i) Likelihood ratio test for the LI effect at $\alpha=0.05$**
Null hypothesis: $H_0: \beta_1=0$\
Alternative hypothesis: $H_A: \beta_1\neq0$\
The test statistic is: $-2(l_0 - l_1)=34.372-26.073=8.299$, which follows a Chi-squared distribution with 1 degree of freedom.\
The p-value is $P(\chi^2\ge 8.299)=0.004 < \alpha$
```{r include=F}
1-pchisq(8.299, df=1)
```
Therefore, we reject the null hypothesis and conclude that we have strong evidence of a positive effect of LI on the cancer remission probability.\

### **(j) 95% likelihood ratio CI for the odds ratio**
The 95% likelihood ratio CI for the log odds ratio, $\beta_1$ is given in the R output, which is $(0.043, 0.285)$.\
Then the 95% likelihood ratio CI for the odds ratio, $e^{\beta_1}$, is $(1.04, 1.33)$. \
*Interpretation:* We are 95% confident that a 1-unit increase in LI has at least 4% increase and at most 33% increase in the odds that a patient achieves cancer remission.\

## **2. Problem 2**

### **(a) Equation for logistic regression model for $\pi$ = probability of a satellite**
$$ln(\frac{\hat{\pi}}{1-\hat{\pi}})=-3.6947+1.8151(weight)$$
```{r}
crabs <- read.table("Crabs.txt", header = T)
crabs.fit <- glm(y ~ weight, family = binomial(link=logit), data = crabs)
crabs.summary <- summary(crabs.fit)
crabs.summary
```

### **(b) Estimated probabilities of having a satellite**
When $weight=1.20$,\
$$\hat{\pi}(1.20)=\frac{e^{-3.6947+1.8151(1.20)}}{1+e^{-3.6947+1.8151(1.20)}} = 0.180$$
When $weight=2.44$,\
$$\hat{\pi}(2.44)=\frac{e^{-3.6947+1.8151(2.44)}}{1+e^{-3.6947+1.8151(2.44)}} = 0.675$$
When $weight=5.20$,\
$$\hat{\pi}(5.20)=\frac{e^{-3.6947+1.8151(5.20)}}{1+e^{-3.6947+1.8151(5.20)}} = 0.997$$
```{r}
b0 <- coef(crabs.fit)[1]
b1 <- coef(crabs.fit)[2]
#predict pi for weight = 1.2
pi.min <- exp(b0+b1*min(crabs$weight)) / (1+ exp(b0+b1*min(crabs$weight)))
pi.min
#predict pi for weight = 2.44
pi.avg <- exp(b0+b1*mean(crabs$weight)) / (1+ exp(b0+b1*mean(crabs$weight)))
pi.avg
#predict pi for weight = 5.2
pi.max <- exp(b0+b1*max(crabs$weight)) / (1+ exp(b0+b1*max(crabs$weight)))
pi.max
```

### **(c) Rate of change in $\hat\pi$ when $x=3.0$**
$$\hat\beta_1\hat\pi(3)(1-\hat\pi(3))=0.229$$
```{r}
pi.3 <- exp(b0+b1*3) / (1+ exp(b0+b1*3))
b1*pi.3*(1-pi.3)
```

### **(d) Weight when $\hat\pi=0.50$**
At the median effective level, the weight is $x = -\frac{\hat\beta_0}{\hat\beta_1}=-\frac{-3.6947}{1.8151}=2.0355$ kg.\
```{r}
-(b0/b1)
```

### **(e) 95% Wald CI for the odds ratio corresponding to a 1-unit increase in weight**
The odds ratio is $e^{\beta_1}$, so we need to construct 95% Wald CI for log odds ratio first, which is $\beta_1$.\
The 95% Wald CI for $beta_1$ is $1.8151 \pm 1.96(0.3767)$ which will be $(1.077, 2.553)$.\
Therefore, the 95% Wald CI for the odds ratio is $(2.935, 12.851)$.\
*Interpretation:* We're 95% confident that a 1-unit increase in weight has at least 193% increase and at most 1185% increase in the odds that a female crab has a satellite.\
```{r}
data.frame("LB"=b1-qnorm(0.975)*crabs.summary$coefficients[2,2], 
           "UB"=b1+qnorm(0.975)*crabs.summary$coefficients[2,2], 
           row.names = "Wald CI for log odds")
data.frame("LB"=exp(b1-qnorm(0.975)*crabs.summary$coefficients[2,2]), 
           "UB"=exp(b1+qnorm(0.975)*crabs.summary$coefficients[2,2]), 
           row.names = "Wald CI for odds")
```

### **(f) Wald test for the weight effect at $\alpha=0.05$**
Null hypothesis: $H_0: \beta_1=0$\
Alternative hypothesis: $H_A: \beta_1\neq0$\
Test statistic is\
$$Z = \frac{\hat\beta_1}{SE(\hat\beta_1)} = \frac{1.815}{0.377}=4.819$$
which follows the standard normal distribution. Thus the p-value is\
$$2P(Z \geq 4.819)= 0.000001445736 < \alpha$$
Therefore, we reject the null hypothesis and conclude that we have strong evidence of a positive effect of weight on the satellite probability of a female crab.\
```{r}
z <- b1/(crabs.summary$coefficients[2,2])
(1-pnorm(z))*2
```

### **(g) Likelihood ratio test for the LI effect at $\alpha=0.05$**
Null hypothesis: $H_0: \beta_1=0$\
Alternative hypothesis: $H_A: \beta_1\neq0$\
The test statistic is: $-2(l_0 - l_1)=225.76-195.74=30.02$, which follows a Chi-squared distribution with 1 degree of freedom.\
The p-value is $P(\chi^2\ge 30.02)=0.00000004273103 < \alpha$
```{r}
chi_sqr <- crabs.summary$null.deviance - crabs.summary$deviance
1-pchisq(chi_sqr, df=1)
```
Therefore, we reject the null hypothesis and conclude that we have strong evidence of a positive effect of weight on the satellite probability of a female crab.\

### **(h) 95% likelihood ratio CI for the odds ratio**
The 95% likelihood ratio CI for the log odds ratio, $\beta_1$ is given in the R output, which is $(1.114, 2.597)$.\
Then the 95% likelihood ratio CI for the odds ratio, $e^{\beta_1}$, is $(3.05, 13.43)$. \
*Interpretation:* We're 95% confident that a 1-unit increase in weight has at least 204.59% increase and at most 1242.75% increase in the odds that a female crab has a satellite.\
```{r}
LR.ci <- confint(crabs.fit)
LR.lo <- data.frame("LB"=LR.ci[2,1], 
           "UB"=LR.ci[2,2], 
           row.names = "Likelihood ratio CI for log odds")
LR.lo
LR.o <- data.frame("LB"=exp(LR.ci[2,1]), 
           "UB"=exp(LR.ci[2,2]), 
           row.names = "Likelihood ratio CI for odds")
LR.o
```